### 2011-08-16 doc/theses/vitasek/chapter-3-mip.tex
Thesis: splitting tex files.

\documentclass[a4paper,12pt,oneside]{report}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
\usepackage{xstring}
\usepackage{wrapfig}
\usepackage{appendix}
\usepackage{nomencl}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{multicol}
\usepackage[pdfborder=0 0 0]{hyperref}
\usepackage{amssymb}
\usepackage{amsthm}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
\def\.{\discretionary{}{}{}}
\newcommand{\heu}[1]{\texttt{\textit{#1}}}
\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
\linespread{1.5}

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\renewcommand{\algorithmicforall}{\textbf{for each}}
\renewcommand{\algorithmiccomment}[1]{\textit{// #1}}

\renewcommand{\nomname}{List of Abbreviations}
\makenomenclature

\theoremstyle{definition}
\newtheorem{define}{Definition}[chapter]

\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{4cm}
\setlength{\evensidemargin}{4cm}
\setlength{\topmargin}{25mm}
\setlength{\footskip}{1cm}
\setlength{\headheight}{0cm}
\setlength{\headsep}{0cm}
\setlength{\marginparwidth}{0cm}
\setlength{\marginparpush}{0cm}
\setlength{\textheight}{23.7cm}
\setlength{\textwidth}{14.5cm}
\let\openright=\clearpage

\def\mfauthor{Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková Ph.D.}
\def\mfplacedate{Praha, 2011}

% Tato makra přesvědčují mírně ošklivým trikem LaTeX, aby hlavičky kapitol
% sázel příčetněji a nevynechával nad nimi spoustu místa. Směle ignorujte.
\makeatletter
\def\@makechapterhead#1{
  {\parindent \z@ \raggedright \normalfont
   \Huge\bfseries \thechapter. #1
   \par\nobreak
   \vskip 20\p@
}}
\def\@makeschapterhead#1{
  {\parindent \z@ \raggedright \normalfont
   \Huge\bfseries #1
   \par\nobreak
   \vskip 20\p@
}}
\makeatother

\def\chapwithtoc#1{
\chapter*{#1}
\addcontentsline{toc}{chapter}{#1}
}

\begin{document}

\lefthyphenmin=2
\righthyphenmin=2

%%% Titulní strana práce ======================================================
\pagestyle{empty}
\begin{center}

\large

Charles University in Prague

\medskip

Faculty of Mathematics and Physics

\vfill

{\bf\Large MASTER THESIS}

\vfill

\centerline{\mbox{\includegraphics[width=60mm]{logo}}}

\vfill
\vspace{5mm}

{\LARGE \mfauthor}

\vspace{15mm}

% exactly as assigned
{\LARGE\bfseries Inference of XML Integrity Constraints}

\vfill

% Název katedry nebo ústavu, kde byla práce oficiálně zadána
% (dle Organizační struktury MFF UK)
Department of Software Engineering

\vfill

\begin{tabular}{rl}

Supervisor of the master thesis: & 	\mfadvisor \\
\noalign{\vspace{2mm}}
Study programme: & Informatika \\
\noalign{\vspace{2mm}}
Specialization: & ISS \\
\end{tabular}

\vfill

% Zde doplňte rok
\mfplacedate

\end{center}

\newpage % ============================================================
%%% Následuje vevázaný list -- kopie podepsaného "Zadání diplomové práce".
%%% Toto zadání NENÍ součástí elektronické verze práce, nescanovat.
\openright

\noindent
[Sample: Here you may thank whoever you wish (the supervisor of the thesis, the consultant,
the person who lent the software, literature, etc.)]

TODO thank jInfer team, Mlynkova, anyone helping create this, colleagues at work...

\newpage % ============================================================
%%% Strana s čestným prohlášením k diplomové práci
\vglue 0pt plus 1fill
\noindent
I declare that I carried out this master thesis independently, and only with the cited
sources, literature and other professional sources.

\medskip\noindent
I understand that my work relates to the rights and obligations under the Act No.
121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
University in Prague has the right to conclude a license agreement on the use of this
work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.

\vspace{10mm}

\hbox{\hbox to 0.5\hsize{%
In ........ date ............
\hss}\hbox to 0.5\hsize{%
signature
\hss}}

\vspace{20mm}

\newpage % ============================================================
%%% Povinná informační strana diplomové práce
\vbox to 0.5\vsize{
\setlength\parindent{0mm}
\setlength\parskip{5mm}

Název práce: Odvozování integritních omezení v XML % TODO how to translate this correctly?

Autor: \mfauthor

Katedra:  Katedra softwarového inženýrství

Vedoucí diplomové práce: \mfadvisor{}, Katedra softwarového inženýrství

Abstrakt: [abstract of 80-200 words in Czech, but not a copy of the assignment of the
master thesis]

Klíčová slova:  XML, ID atributy, odvozování

\vss}\nobreak\vbox to 0.49\vsize{
\setlength\parindent{0mm}
\setlength\parskip{5mm}

Title:  Inference of XML Integrity Constraints

Author: \mfauthor

Department: Department of Software Engineering

Supervisor: \mfadvisor{}, Department of Software Engineering

Abstract:  [abstract of 80-200 words in English, but not a copy of the assignment of
the master thesis]

Keywords: XML, ID attributes, inference

\vss}

\newpage

%%% Strana s automaticky generovaným obsahem diplomové práce. U matematických
%%% prací je přípustné, aby seznam tabulek a zkratek, existují-li, byl umístěn
%%% na začátku práce, místo na jejím konci.

\openright
\pagestyle{plain}
\setcounter{page}{1}
\tableofcontents

% =============== TEXT ================
\newpage

% TODO what style of quotation marks to use? 
% TODO how to format text? is textit / texttt / ... OK?

\chapwithtoc{Preface}

Along with technologies like JSON, SQL/noSQL databases and bla, XML is one of the leading formats for storing structured data. However, even though languages such as DTD and XML Schema to describe XML structure already exist since a long time, most of the documents use outdated or no schema at all (link Vosta's Even ant can create...). To tackle this problem one may employ reverse-engineering techniques to infer the schema from existing documents, such as those described in A, B, C, jInfer. But the schema is not the only constraint that can be imposed on an XML document: the concept of \textit{keys} and \textit{foreign keys}, well known from the relational database world, applies here as well. One could go even further and try to find even more sophisticated relations in the data, such as \textit{functional dependencies} (link Sviro).

This work will be building upon the achievements of jInfer schema inference framework (TODO link Anti's improvements in schema inference), expanding its possibilities in the field of search for \textit{key-} and \textit{foreign key-}like structures in existing XML documents.

TODO Integrity constraints can be keys (with ID attributes as a "sub-group"), FKs, functional dependencies (quote Sviro), etc.
We will focus on the first kind - ID and IDREF attributes.

TODO argument: test data with DTD (and thus possibly ID/IDREF) is more common, we can have better test sets.

\section{Structure of the thesis}

The thesis will be structured as follows.

First, we will introduce a few notions required throughout the work, such as XML tree, ID attributes, ID sets and keys for XML.

Secondly, we will review approaches to ID attribute and XML keys search from previous articles on this topic.

This will lead us to the NP-complete problem of maximal independent set (IS)\nomenclature{IS}{Independent Set}, where we will inspect approaches for solving it.

We will discuss a closely related Mixed Integer Problem (MIP)\nomenclature{MIP}{Mixed Integer Problem} and prove their "equality".

Afterwards, we will show how to use an external MIP solver and various heuristics to tackle this problem.

An extension to jInfer for finding ID attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in the closing chapters.

\section{Conventions}

TODO list conventions - how we write code related stuff, module names, heu names, etc

\chapter{Definitions}

\begin{itemize}
	\item XML tree - from FIDAX or wherever
	\item Element, Attribute
	\item ID, IDREF, IDREFS attribute according to specification
	\item XSD keys (compare them to ID attributes)
	\item Key according to Keys for XML \cite{keX}
	\item Attribute mapping \nomenclature{AM}{Attribute Mapping}
	\item ID set
  \item Weight: support, coverage
\end{itemize}

\chapter{Research}

According to the article \cite{fidax}, ...

TODO talk about FIDAX
To the best of our knowledge, there are no other articles dealing with this problem.

TODO talk about Fajt - but that's different, that's keys
TODO add citation for Fajt

\section{Independent Set}

TODO define IS + MIS rigorously

TODO other approaches to max IS

@@

\begin{algorithm}
\caption{FIDAX CH}
\label{listing-ch-fidax}
\begin{algorithmic}
\REQUIRE $C$ list of candidate AMs
\ENSURE a feasible solution
\STATE $C' \gets C$ sorted by decreasing size
\STATE Compute the weight $w(m)$ of each $m$ in $C$
\FORALL{$t$ in $\Sigma^E$}
  \STATE Let $m$ be a \textbf{highest-weight} mapping of type $t$ in $C'$
  \STATE Remove from $C'$ all mappings of type $t$ except $m$
\ENDFOR
\FORALL{$m$ in $C'$}
  \STATE $S \gets$ all mappings in $C$ whose images intersect $\iota(m)$
  \IF{$w(m) > \sum_{p \in S} w(p)$}
    \STATE remove all $p \in S$ from $C'$
  \ELSE
    \STATE remove $m$ from $C'$
  \ENDIF
\ENDFOR
\RETURN $C'$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Random CH}
\label{listing-ch-random}
\begin{algorithmic}
\REQUIRE $N$ required size of pool
\REQUIRE $C$ list of candidate AMs
\ENSURE pool of $N$ feasible solutions
\STATE $r \gets $ empty pool
\FOR{$i = 1 \to N$} 
  \STATE \COMMENT{create 1 solution}
  \STATE $s \gets $ empty solution
  \WHILE{$s$ is a feasible ID set}
    \STATE $a \gets $ pick at random from $C \backslash S$
    \STATE $s \gets s \cup a$
  \ENDWHILE
  \STATE $r \gets r \cup s$
\ENDFOR
\RETURN $r$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Fuzzy CH}
\label{listing-ch-fuzzy}
\begin{algorithmic}
\REQUIRE $N$ required size of pool
\REQUIRE $C$ list of candidate AMs
\ENSURE pool of $N$ feasible solutions
\STATE $r \gets $ empty pool
\FOR{$i = 1 \to N$} 
  \STATE \COMMENT{create 1 solution}
  \STATE $s \gets $ empty solution
  \STATE $C' \gets C$

@@

\begin{algorithm}
\caption{Incremental CH}
\label{listing-ch-incremental}
\begin{algorithmic}
\REQUIRE $C$ list of candidate AMs
\ENSURE a feasible solution
\STATE $C' \gets $ sort $C$ by decreasing weight
\STATE $s \gets $ empty solution
\FORALL{$c \in C'$}
  \IF{$s \cup c$ is a feasible ID set}
    \STATE $s \gets s + c$
  \ENDIF
\ENDFOR
\RETURN $s$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Removal CH}
\label{listing-ch-removal}
\begin{algorithmic}
\REQUIRE $C$ list of candidate AMs
\ENSURE a feasible solution
\STATE $C' \gets $ sort $C$ by increasing weight
\STATE $s \gets C'$
\FORALL{$c \in s$}
  \IF{$s$ is a feasible ID set}
    \RETURN $s$
  \ENDIF
  \STATE $s \gets s \backslash c$
\ENDFOR
\end{algorithmic}
\end{algorithm}

@@

@@

\begin{algorithm}
\caption{Identity IH}
\label{listing-ih-identity}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\ENSURE the same pool of feasible solutions
\RETURN $FP$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Remove Worst IH}
\label{listing-ih-removeworst}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\ENSURE pool of feasible solutions
\STATE $s_{min} \gets $ solution with the lowest weight $\in FP$
\RETURN $FP \backslash s_{min}$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Random Remove IH}
\label{listing-ih-randomremove}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\REQUIRE $k \in (0,1)$ ratio of AMs to remove from each $s \in FP$
\ENSURE pool of feasible solutions
\FORALL{$s \in FP$}
  \STATE $K \gets k * |s|$
  \STATE remove $K$ random AMs from $s$
\ENDFOR
\RETURN $FP$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Hungry IH}
\label{listing-ih-hungry}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\REQUIRE $C$ list of candidate AMs
\ENSURE pool of feasible solutions
\FORALL{$s \in FP$}
  \STATE \COMMENT {improve a single solution}
  \STATE $C' \gets C \backslash s$
  \STATE $C' \gets C'$ sorted by decreasing weight
  \FORALL{$c \in C'$}
    \IF{$s \cup c$ is a feasible ID set}
      \STATE $s \gets s \cup c$
    \ENDIF
  \ENDFOR
\ENDFOR
\RETURN $FP$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Mutation IH}
\label{listing-ih-mutation}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\REQUIRE $k$ ratio of AMs to fix
\ENSURE pool of feasible solutions
\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
\STATE $K \gets k * |incumbent|$
\STATE fix $K$ random AMs from $incumbent$ in GLPK problem formulation
\STATE $improved \gets $ run GLPK
\RETURN $FP \cup improved$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Crossover IH}
\label{listing-ih-crossover}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\REQUIRE $k$ ratio of solutions to look for commonalities in
\ENSURE pool of feasible solutions
\STATE $K \gets k * |FP|$
\STATE $FP' \gets K$ random solutions $\in FP$
\STATE $am \gets$ AMs found in all solutions $\in FP'$
\STATE fix $am$ in GLPK problem formulation
\STATE $improved \gets $ run GLPK
\RETURN $FP \cup improved$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{Local Branching IH}
\label{listing-ih-localbranching}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\REQUIRE $k$ ratio of total AM count to bound the Hamming distance to
\ENSURE pool of feasible solutions
\STATE $K \gets k * |$total AM count$|$
\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
\STATE add max Hamming distance requirement to GLPK problem formulation
\STATE $improved \gets $ run GLPK
\RETURN $FP \cup improved$
\end{algorithmic}
\end{algorithm}

@@

\chapter{Experiments}

At this point of the thesis the reader should be already familiar with the notions we have introduced: the problem of finding the optimal ID set (with respect to some \textit{weight}), that it is directly related to the NP-complete problem of finding the maximal weighted independent set, that this can be solved using the MIP approach, and that we can try to do better than just let the solver work: by employing various heuristics.

Now it is the time to move our ideas into reality and test their feasibility. But before we start talking about the experiments themselves, we should try to formulate our aim, what we will be trying to establish.\\

First of all, we would like to get an idea of how the whole system and its components behave. We would like to see the changes introduced by modifying some of the key parameters, while keeping the others fixed. Even though we shan't hope for them to be orthogonal, we might at least isolate some of the parameters that are less important to the overall behavior. Preferably, in the end we should have at least some intuition into what will happen if we try X Y Z, before trying it.

Second, as we will be introducing a few broad classes of XML data to run ID set search on, we would like to find out which types are best handled by which configurations and settings. This will allow us to pick the right heuristic once we see new data.

Also, we will try to evaluate the system performance in terms of the speed of finding good heuristic results. We will try to find tweaks to make the whole process as fast as reasonably possible.

And in the end, we should be able to formulate some kind of general recommendation in form ``if you see this kind of data, do that".\\

This chapter will be structured in the following way: first we will discuss the experimental data we used, then the methodology used in conducting the experiments, followed by the actual list of experiments with their full description and results, and in the end we shall draw some final conclusions.

\section{Experimental Data}

Let us now talk about the XML data we will be using to conduct out experiments. We are using XML documents of three categories:

\begin{itemize}
	\item Realistic 
	\item Realistic with artificial (converted) attributes
	\item Artificial
\end{itemize}

A short reasoning for this choice: realistic - of course, we want to see the performance in cases taken from the real world. The problem with realistic data is that sometimes, interesting values (that might or might not contain IDs) are stored as text nodes % TODO is this name consistent with what I've been defining?
instead of attributes. We might try to look at such data, convert some of these ``suspicious" values to attributes (e.g. using a smart XSL transformation), let our heuristics find the ID sets, and then translate them back to XML keys (see \ref{realistic-converted} for details).
And finally, we create completely artificial data to create inputs that will really put our heuristics to the test. This is because the realistic data often prove to be quite easy to solve - the list of candidate AMs is usually too short to be hard to solved to optimality.\\

To understand these data sets, we will talk a little about their origin and \textit{graph representation}. As mentioned earlier, % TODO link
the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set on a graph. Therefore, it might be interesting to actually see the graphs of these data sets and have some kind of numbers associated with them.

The former will be achieved with the help of GraphViz tool, % TODO link
where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of their images (and thus cannot be in the same ID set together). Thus solving the maximal weighted IS on these graphs will be equivalent to solving our problem of optimal ID set.

The latter will come in form of tables containing information regarding the data sets, such their size, known optimum % TODO footnote explaining that we got the optimum by running GLPK w/o time limit
for $\alpha = \beta = 1$ and the numbers of vertices and edges in aforementioned graphs.

\subsection{Realistic data}

From 3 different sources we collected 6 different data sets, called \jmodule{OVA1} - \jmodule{OVA3}, \jmodule{XMA-c}, \jmodule{XMA-p} and \jmodule{XMD}. Their summary is the Table \ref{table-experiments-data-realistic}, their graphs can be seen in Figure \ref{image-experiments-data-realistic}. Because the legal status of disclosing these data sets is unclear, we will refrain from identifying them beyond these artificial identifiers. Neither will they be included on the DVD distributed with this thesis.

TODO their DTD/XSD - do we have ID attributes?

TODO verify they satisfy their schemas. Converted too.

\begin{table}
  \caption{List of realistic test data files}
  \bigskip
  \label{table-experiments-data-realistic}
  \centering
  \begin{tabular}{l | r | c | c | l}
	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
	\hline
	OVA1  & 4.5      & 29 & 43 & 0.45588235294117635 \\
	OVA2  & 11.9     & 23 & 36 & 0.1634615384615385  \\ 
	OVA3  & 237.6    & 31 & 47 & 0.25537156151635415 \\
	XMA-c & 1 807.7  & 1  & 0  & 0.7546666666666667  \\
	XMA-p & 13 748.3 & 1  & 0  & 0.2019306150568969  \\
	XMD   & 1 743.0  & 17 & 15 & 0.09786094165493507 \\
  \end{tabular}
\end{table}

\begin{figure}
  \caption{Realistic data}
  \label{image-experiments-data-realistic}
  \centering
    \subfigure[OVA1]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA1}}
    \subfigure[OVA2]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA2}}
    \subfigure[OVA3]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA3}}
    \subfigure[XMA-c]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-c}}
    \subfigure[XMA-p]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-p}}
    \subfigure[XMD]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMD}}
\end{figure}

To interpret the data a little: \jmodule{OVA*} sets have quite interesting and perhaps a bit challenging graphs, but are not too big. We can consider them to be the ``typical" representants.

On the other hand, the \jmodule{XMA-*} sets are quite huge, but trivial: their only candidate AM will just get picked and that will be the end of the heuristic. We will see the performance of the other components of the whole system, dealing with huge data that are rather simple in the end.

Finally, the \jmodule{XMD} set is quite big and at the same time has non-trivial graph representation. We should see a performance more balanced between processing and finding the ID set here.

\subsection{Realistic data with artificial attributes}
\label{realistic-converted}

We found 2 data sets to convert, \jmodule{MSH} and \jmodule{NTH}. Unfortunately, the same problem with disclosure as in the previous case applies here. None of these sets had any attributes before the conversion. Their summary is the Table \ref{table-experiments-data-converted}, their graphs can be seen in Figure \ref{image-experiments-data-converted}.

TODO schemas?

To address the conversion: in the case of \jmodule{MSH} we found 2 elements with values looking suspiciously like a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation. In the case of \jmodule{NTH} we converted all the values in sub-elements of the record elements to be the attributes of the records.

Why is this useful at all? Recall TODO link, where we said that ID attributes are a special case of XML keys. We can use this approach then to find XML keys: convert some suspicious data into attributes, find the optimal ID set and then create XML key based on this ID set.

\begin{table}
  \caption{List of realistic test data files with converted attributes}
  \bigskip
  \label{table-experiments-data-converted}
  \centering
  \begin{tabular}{l | r | c | c | l}
	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
	\hline
	MSH  & 3 100.5 & 1 & 0 & 0.5416472778036296 \\
	NTH  & 2 523.5 & 5 & 7 & 0.057918595422124436 \\ 
  \end{tabular}
\end{table}

\begin{figure}
  \caption{Realistic data with converted attributes}
  \label{image-experiments-data-converted}
  \centering
    \subfigure[MSH]{\includegraphics[width=0.08\textwidth]{images/experiments/data/converted/MSH}}
    \subfigure[NTH]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/NTH}}
\end{figure}

Again to do a little interpretation: in the case of \jmodule{MSH} we created 2 attributes, of which only one constituted a candidate AM. This is then the case similar to \jmodule{XMA-*} sets: quite large data, yet only one trivial ID attribute to be found.

In the case of \jmodule{NTH} we introduced more attributes, 8 to be precise. Out of them 5 proved to be candidate AMs, with 7 edges constraining them. This means we have quite a big set with relatively simple work to be done by the heuristics. This is about the change.

\subsection{Artificial data}

As soon as we started experimenting with the data coming from the real world, it was obvious that they present no real challenge to our heuristics. After we build the model, we get the most complex graphs of some 31 vertices and 47 edges (see Table \ref{table-experiments-data-realistic}). We needed to create data that would really put pressure on our heuristics, put them to a real test. Our solutions is to approach from the other side: in the end, we will be solving the equivalent of IS problem on a graph created from XML data. Why not create the XML data to contain a more complex graph in the first place? What if we could specify how many vertices and edges we want, and get the according XML?

This is actually simple. Consider the following excerpt from an XML file:

\begin{scriptsize}
\begin{verbatim}
<graph>
  <vertex0 attr="-2968876296119015800"/>
  <vertex1 attr="1729745997570096518"/>
  <vertex2 attr="-9020549659620928934"/>
  ...
  <vertex99 attr="-7545982394508643394"/>

  <vertex82 attr="0"/><vertex21 attr="0"/>
  <vertex64 attr="1"/><vertex21 attr="1"/>
  <vertex44 attr="2"/><vertex2 attr="2"/>
  ...
  <vertex96 attr="99"/><vertex40 attr="99"/>
</graph>
\end{verbatim}
\end{scriptsize}

We want to create a graph with around $v$ vertices and $e$ edges. First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}. To constitute an AM, they need an attribute \texttt{attr}, but with large enough random values, so that these values won't conflict with any others. Second, for each of the $e$ edges we pick two \texttt{vertex*} elements at random, and give them the same value of their \texttt{attr}. This will ensure they cannot share the same ID set, thus effectively creating the edge in the graph representation.

The pseudocode for this is in listing \ref{listing-random-data}.

\begin{algorithm}
\caption{Random XML data creation}
\label{listing-random-data}
\begin{algorithmic}
\REQUIRE $v$ requested number of vertices
\REQUIRE $e$ requested number of edges
\ENSURE XML file content
\PRINT \texttt{<graph>}
\FOR{$i = 1 \to |V|$}
	\STATE $R \gets RANDOM$
	\PRINT \texttt{<vertex\textit{i} attr="\textit{R}">}
\ENDFOR
\FOR{$i = 1 \to |E|$}
	\STATE $v1 \gets RANDOM(|V|)$
	\STATE $v2 \gets RANDOM(|V|)$
	\PRINT \texttt{<vertex\textit{v1} attr="\textit{i}"> <vertex\textit{v2} attr="\textit{i}">}
\ENDFOR
\PRINT \texttt{</graph>}
\RETURN
\end{algorithmic}
\end{algorithm}

With this process in place we can create as much data as we wish with any combination of $v$ and $e$ we want.

There is one interesting characteristic that can describe random graphs like this, and that is the \textit{density}. This can be defined in various ways, we will use two different interpretations for a while. The first is $\frac{|E|}{|V|}$, that is, how many edges are there for one vertex (multiplied by 2 we'd get the average degree of the vertices). The second, perhaps more interesting is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$. This is the density as the ratio of edges that are to all edges that could be in a complete graph with $|V|$ vertices.

We have created 3 sets to use in experiments along with the realistic and converted sets, these are called \jmodule{100-100}, \jmodule{100-200} and \jmodule{100-1000}. Note that the name is always in the form $v-e$.\\

All of the experimental data sets mentioned so far, realistic, converted and artificial alike will be referred to as \textit{official test data (sets)}.\\

Also, we will need data of comparably similar characteristics but varying size, to study the effects of size on the run times of experiments. For this reason we created 11 more sets, from \jmodule{0-0} as the trivial one to \jmodule{100-500} as the biggest one. From now on, these will be referred to as \textit{sized test data (sets)}. We wanted to keep density the same among these sets, and we picked the $\frac{|E|}{E_{max}}$ density interpretation for this.

TODO this is on the DVD.

The summary is the Table \ref{table-experiments-data-artificial} and Table \ref{table-experiments-data-artificial-size}, these tables contain 2 new columns: values of density in both interpretations we introduced. Some graphs can be seen in Figure \ref{image-experiments-data-artificial}.

While studying the tables you might notice that the actual numbers $|V|$ and $|E|$ do not match to the $v$ and $e$ in the names of the sets. This is because how the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set (proof is homework). % TODO too much?:)
Because of the so-called Birthday paradox, % TODO link
this will happen more with higher $e$.

>>>>

@@

\begin{algorithm}
\caption{\jmodule{FIDAX} CH}
\label{listing-ch-fidax}
\begin{algorithmic}
\REQUIRE $C$ list of candidate AMs
\ENSURE a feasible solution
\STATE $C' \gets C$ sorted by decreasing size
\STATE Compute the weight $w(m)$ of each $m$ in $C$
\FORALL{$t$ in $\Sigma^E$}
  \STATE Let $m$ be a \textbf{highest-weight} mapping of type $t$ in $C'$
  \STATE Remove from $C'$ all mappings of type $t$ except $m$
\ENDFOR
\FORALL{$m$ in $C'$}
  \STATE $S \gets$ all mappings in $C$ whose images intersect $\iota(m)$
  \IF{$w(m) > \sum_{p \in S} w(p)$}
    \STATE remove all $p \in S$ from $C'$
  \ELSE
    \STATE remove $m$ from $C'$
  \ENDIF
\ENDFOR
\RETURN $C'$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{\jmodule{Random} CH}
\label{listing-ch-random}
\begin{algorithmic}
\REQUIRE $N$ required size of pool
\REQUIRE $C$ list of candidate AMs
\ENSURE pool of $N$ feasible solutions
\STATE $r \gets $ empty pool
\FOR{$i = 1 \to N$}
  \STATE \COMMENT{create 1 solution}
  \STATE $s \gets $ empty solution
  \WHILE{$s$ is a feasible ID set}
    \STATE $a \gets $ pick at random from $C \backslash S$
    \STATE $s \gets s \cup a$
  \ENDWHILE
  \STATE $r \gets r \cup s$
\ENDFOR
\RETURN $r$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{\jmodule{Fuzzy} CH}
\label{listing-ch-fuzzy}
\begin{algorithmic}
\REQUIRE $N$ required size of pool
\REQUIRE $C$ list of candidate AMs
\ENSURE pool of $N$ feasible solutions
\STATE $r \gets $ empty pool
\FOR{$i = 1 \to N$}
  \STATE \COMMENT{create 1 solution}
  \STATE $s \gets $ empty solution
  \STATE $C' \gets C$

@@

\begin{algorithm}
\caption{\jmodule{Incremental} CH}
\label{listing-ch-incremental}
\begin{algorithmic}
\REQUIRE $C$ list of candidate AMs
\ENSURE a feasible solution
\STATE $C' \gets $ sort $C$ by decreasing weight
\STATE $s \gets $ empty solution
\FORALL{$c \in C'$}
  \IF{$s \cup c$ is a feasible ID set}
    \STATE $s \gets s + c$
  \ENDIF
\ENDFOR
\RETURN $s$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{\jmodule{Removal} CH}
\label{listing-ch-removal}
\begin{algorithmic}
\REQUIRE $C$ list of candidate AMs
\ENSURE a feasible solution
\STATE $C' \gets $ sort $C$ by increasing weight
\STATE $s \gets C'$
\FORALL{$c \in s$}
  \IF{$s$ is a feasible ID set}
    \RETURN $s$
  \ENDIF
  \STATE $s \gets s \backslash c$
\ENDFOR
\end{algorithmic}
\end{algorithm}

@@

TODO we shuffle AMs - we get different runs - pool is possible

@@

\begin{algorithm}
\caption{\jmodule{Identity} IH}
\label{listing-ih-identity}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\ENSURE the same pool of feasible solutions
\RETURN $FP$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{\jmodule{Remove Worst} IH}
\label{listing-ih-removeworst}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\ENSURE pool of feasible solutions
\STATE $s_{min} \gets $ solution with the lowest weight $\in FP$
\RETURN $FP \backslash s_{min}$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{\jmodule{Random Remove} IH}
\label{listing-ih-randomremove}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\REQUIRE $k \in (0,1)$ ratio of AMs to remove from each $s \in FP$
\ENSURE pool of feasible solutions
\FORALL{$s \in FP$}
  \STATE $K \gets k * |s|$
  \STATE remove $K$ random AMs from $s$
\ENDFOR
\RETURN $FP$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{\jmodule{Hungry} IH}
\label{listing-ih-hungry}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\REQUIRE $C$ list of candidate AMs
\ENSURE pool of feasible solutions
\FORALL{$s \in FP$}
  \STATE \COMMENT {improve a single solution}
  \STATE $C' \gets C \backslash s$
  \STATE $C' \gets C'$ sorted by decreasing weight
  \FORALL{$c \in C'$}
    \IF{$s \cup c$ is a feasible ID set}
      \STATE $s \gets s \cup c$
    \ENDIF
  \ENDFOR
\ENDFOR
\RETURN $FP$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{\jmodule{Mutation} IH}
\label{listing-ih-mutation}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\REQUIRE $k$ ratio of AMs to fix
\ENSURE pool of feasible solutions
\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
\STATE $K \gets k * |incumbent|$
\STATE fix $K$ random AMs from $incumbent$ in GLPK problem formulation
\STATE $improved \gets $ run GLPK
\RETURN $FP \cup improved$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{\jmodule{Crossover} IH}
\label{listing-ih-crossover}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\REQUIRE $k$ ratio of solutions to look for commonalities in
\ENSURE pool of feasible solutions
\STATE $K \gets k * |FP|$
\STATE $FP' \gets K$ random solutions $\in FP$
\STATE $am \gets$ AMs found in all solutions $\in FP'$
\STATE fix $am$ in GLPK problem formulation
\STATE $improved \gets $ run GLPK
\RETURN $FP \cup improved$
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{\jmodule{Local Branching} IH}
\label{listing-ih-localbranching}
\begin{algorithmic}
\REQUIRE $FP$ pool of feasible solutions
\REQUIRE $k$ ratio of total AM count to bound the Hamming distance to
\ENSURE pool of feasible solutions
\STATE $K \gets k * |$total AM count$|$
\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
\STATE add max Hamming distance requirement to GLPK problem formulation
\STATE $improved \gets $ run GLPK
\RETURN $FP \cup improved$
\end{algorithmic}
\end{algorithm}

@@
